Visual inference for networks

### 2016-02-17 paper/network-inference.Rnw
working on sims

ggplot(data=friend.data.net.all, aes(from_id = from, to_id = to)) +
  geom_net(aes(colour = as.factor(drink)), size = .75, vjust = .5,
           fontsize = 6, curvature = .25, ealpha = .6, directed = T,
           arrowsize = .5, arrowgap = .01) + theme_net() + 
  scale_color_brewer(palette = "YlOrRd", name = "Drinking Frequency", labels = c("never", "about 2x/year", "about 1x/month", "about 1x/week", "More than 1x/week")) + 
  facet_wrap(~year, nrow = 1, labeller="label_both") +
  theme(legend.position="bottom") +
  guides(col = guide_legend(nrow = 2, override.aes = list(size=3.5)))
@
\caption{\label{fig:friendship_waves}\st{The friendship network of 50 girls in the \emph{Teenage Friends and Lifestyle Study} over three time periods. The nodes, which represent the girls in the study, are colored by their level of alcohol use.}}
\end{figure}
Several actor-oriented longitudinal network models can be fit to this data in RSiena. In order to determine the importance of certain actor-level variable effects included in the models, we simulated networks from different fitted models in Siena, and placed them in lineups to determine if there is any difference between the models.

\par In the fitting of models to the Friends data, there are four effects that are automatically included in model estimation. Two of these are rate parameters for moving from time point one to two and from time point two to three. The other two are the outdegree and the reciprocity effects. The outdegree is just an edge count from each node, and the reciprocity is the count of reciprocated ties from each node. XXX insert more about the math! XXX The documentation and code examples from the RSiena webpage include models to fit to the Friends' data. One of these includes the effects covariate-alter, covariate-ego, and covariate-alter X ego, which are measures of the behavior of a node and its neighbors. XXX insert more math XXX. The RSiena and RSienaTest packages include functions to test the significance of fitted parameters via t-tests, wald tests, and score-type tests. I used the wald tests and the score-type test to assess the significance of the parameters in this friends model with three covariate effects. The tests found the two rate parameters and the outdegree and reciprocity parameters were significant, but of the covariate effects, only the covariate-alter X ego effect was significant. So, we try to determine if there was a difference visually between the networks with just the four network level variables included and the networks with the significant covariate parameter included.
}

>>>>

ggplot(data=friend.data.net.all, aes(from_id = from, to_id = to)) +
  geom_net(aes(colour = as.factor(drink)), size = .75, vjust = .5,
           fontsize = 6, curvature = .25, ealpha = .6, directed = T,
           arrowsize = .5, arrowgap = .01) + theme_net() + 
  scale_color_brewer(palette = "YlOrRd", name = "Drinking Frequency", labels = c("never", "about 2x/year", "about 1x/month", "about 1x/week", "More than 1x/week")) + 
  facet_wrap(~year, nrow = 1, labeller="label_both") +
  theme(legend.position="bottom") +
  guides(col = guide_legend(nrow = 2, override.aes = list(size=3.5)))
@
\caption{\label{fig:friendship_waves}\st{The friendship network of 50 girls in the \emph{Teenage Friends and Lifestyle Study} over three time periods. The nodes, which represent the girls in the study, are colored by their level of alcohol use.}}
\end{figure}

Several actor-oriented longitudinal network models can be fit to this data in RSiena. In order to determine the importance of certain actor-level variable effects included in the models, we simulated networks from different fitted models in Siena, and placed them in lineups to determine if there is any difference between the models.

\par In the fitting of models to the Friends data, there are four effects that are automatically included in model estimation. Two of these are rate parameters for moving from time point one to two and from time point two to three. The other two are the outdegree and the reciprocity effects. The outdegree is just an edge count from each node, and the reciprocity is the count of reciprocated ties from each node. XXX insert more about the math! XXX The documentation and code examples from the RSiena webpage include models to fit to the Friends' data. One of these includes the effects covariate-alter, covariate-ego, and covariate-alter X ego, which are measures of the behavior of a node and its neighbors. XXX insert more math XXX. The RSiena and RSienaTest packages include functions to test the significance of fitted parameters via t-tests, wald tests, and score-type tests. I used the wald tests and the score-type test to assess the significance of the parameters in this friends model with three covariate effects. The tests found the two rate parameters and the outdegree and reciprocity parameters were significant, but of the covariate effects, only the covariate-alter X ego effect was significant. So, we try to determine if there was a difference visually between the networks with just the four network level variables included and the networks with the significant covariate parameter included.

\par To simulate networks from these two models, we first fit RSiena models with the default effects and with the default effects plus the significant actor-level covariate variable. After obtaining the estimates of the effects, we simulated 
}
